---
title: "U.S. Incarceration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('../source/analysis.R')
```
## Introduction
In this report, I will be discussing the issue of disproportionate incarceration
in the United States. I will be using a data set from Vera Institute that
contains data about jail and prison populations of each county in the U.S.,
broken down by groups, such as race and sex. The data set uses 5 racial groups:
Black, White, Asian American/Pacific Islander, Native, and Latinx. For this 
report, I will be focusing on the statistics regarding Black Americans. 
`r percent_black_pop`% of Americans are Black, yet `r percent_black_prison_pop`%
of the prison population is Black, an extremely disproportionate amount. This
disproportionality is not seen in other racial groups. For example,
`r percent_latinx_pop`% of Americans are Latinx, and 
`r percent_latinx_prison_pop`% of the prison population is Latinx, which is much
more proportionate. This issue is seen even greater on smaller scales, in fact,
`r county_max_black_prison_percent` has `r max_county_black_prison_percent`%
of its Black population imprisoned.

## Incarceration Over Time
```{r time_plot, echo = FALSE, fig.show = 'hold'}
plot(time_plot)
```
<br />
I chose to graph total prison population by race over time to see how the issue
of disproportionate incarceration has changed over time. This graph shows that 
Black Americans have historically always had the largest incarcerated population
despite making up a fraction of the total U.S.population. Latinx Americans also 
a large portion of the prison population, while Asian American/Pacific Islanders
and Indigenous Americans contribute a very small amount. White Americans have 
historically been the second largest racial group in prison; however, it is 
important to keep in mind they make up the majority of the total population. 
It also shows an interesting pattern: White, Black, and Latinx prison population
often increase and decrease at similar times. This could be due to general 
legislation changes that effect these groups similarly.

## Black Population vs. Black Imprisonment
```{r comp_plot, echo = FALSE, message = FALSE, warning = FALSE, fig.show = 'hold'}
plot(comp_plot)
```
<br />
I chose to create a scatter plot of the percent of the population that is Black
versus the percent of the prison population that is black per county in order to
see if this issue is more prevalent in areas where Black Americans are more of a
minority group. I chose to use a log scale for both axes to help visualize the
large range of rates, and I added a line of best fit to help make out a trend.
The graph shows that the rate of incarceration is inversely proportionate to the 
population rate, meaning Black Americans are imprisoned more in counties in
which they are more in the minority.

## Map
```{r map, echo = FALSE, fig.show = 'hold'}
plot(map)
```
<br />
I chose to shade each state based off of the percent of the Black population
that is in prison, as I wanted to see if geography is a factor in this issue,
and only looking at the total number of imprisoned Black Americans would not 
take into account the population of the state. The map shows that Black 
incarceration vary a lot between states, but it is difficult to see if there are
more general geographic trends. It seems as if the rate is generally higher
in the East compared to the West, but there are many states in both areas with
a rate near 0.